### 2022-06-01 ms.Rnw
a few minor changes

\fontsize{12}{14pt plus.8pt minus .6pt}\selectfont \vspace{0.8pc}
\centerline{\large\bf An iterated block particle filter for inference on coupled dynamic}
\vspace{2pt} 
\centerline{\large\bf systems with shared and unit-specific parameters}
\vspace{.4cm} 
\centerline{Edward. L. Ionides, Ning Ning and Jesse Wheeler}
\vspace{.4cm} 
\centerline{University of Michigan, Department of Statistics,}

@@

The simulated model is drawn from model (A), which is nested within model (B) but not within (C).
For models B and C, with all parameters are unit-specific, we must estimate $20\times 13=260$ parameters.
This greatly exceeds the 7 shared parameters fitted by \citet{ionides21} for a measles metapopulation model using a large computational effort.
An iterated guided intermediate resampling filter (IGIRF) algorithm was used to fit 9 shared parameters and 3 unit-specific initial value parameters in a measles metapopulation model \citep{park20}.
That analysis used a customized treatment for the unit-specific initial value parameters and did not attempt to estimate other unit-specific parameters.
IGIRF is sensitive to the choice of the guide function, and the model-specific implementation used by \citet{park20} performed better than the generic implementation of IGIRF which is currently available in \code{spatPomp}.
To our knowledge, {\ibpf} is therefore advancing the current limits for scalability of likelihood-based inference for metapopulation dynamics.

@@

We compartmentalize the population of each city into susceptible ($S$), exposed ($E$), infectious ($I$), and recovered/removed ($R$) categories.
The numbers of individuals in each compartment for city $\unit$ at time $t$ are denoted by integer-valued random variables $S_\unit(t)$, $E_\unit(t)$, $I_\unit(t)$, and $R_\unit(t)$.
The population dynamics are written in terms of counting processes $N_{AB,\unit}(t)$ enumerating cumulative transitions in city $\unit$, up to time $t$, between compartments $A$ and $B$.

@@

We enumerate the $U=20$ cities studied by \citet{he10} in decreasing size, so that $\unit=1$ corresponds to London.

@@

The number of recovered individuals $R_\unit(t)$ in city $\unit$ is therefore defined implicitly.

@@

In \eqref{eq:dEdt}, $\alpha$ is a mixing exponent modeling inhomogeneous contact rates within a city, and $\iota$ models immigration of infected individuals which is appropriate when analyzing a subset of cities that cannot be treated as a closed system.
The number of travelers from city $\unit$ to $\altUnit$ is denoted by $v_{\unit\altUnit}$. 

@@

where $\dist(\unit,\altUnit)$ denotes the distance between city $\unit$ and city $\altUnit$, $P_\unit$ is the average population for city $\unit$ across time, $\bar{P}$ is the average population across cities, and $\overline{\dist}$ is the average distance between a randomly chosen pair of cities.
Here, we model $v_{\unit\altUnit}$ as fixed through time and symmetric between any two arbitrary cities, though a natural extension would allow for temporal variation and asymmetric movement between two cities.

@@

\noindent This work was supported by National Science Foundation grant DMS-1761603 and DMS 1646108.
We acknowledge an anonymous referee from a previous article who recommended that we should try implementing a block particle filter.

>>>>

\fontsize{12}{14pt plus.8pt minus .6pt}\selectfont \vspace{0.8pc}
\centerline{\large\bf An iterated block particle filter for inference on coupled dynamic}
\vspace{2pt} 
\centerline{\large\bf systems with shared and unit-specific parameters}
\vspace{.4cm} 
\centerline{Edward. L. Ionides, Ning Ning and Jesse Wheeler}
\vspace{.4cm} 
\centerline{University of Michigan, Department of Statistics}

@@

The simulated model is drawn from model (A), which is nested within model (B) but not within (C).
For models B and C, with all parameters are unit-specific, we must estimate $20\times 13=260$ parameters.
This greatly exceeds the 7 shared parameters fitted by \citet{ionides21} for a measles metapopulation model using a large computational effort.
An iterated guided intermediate resampling filter (IGIRF) algorithm was used to fit 9 shared parameters and 3 unit-specific initial value parameters in a measles metapopulation model \citep{park20}.
That analysis used a customized treatment for the unit-specific initial value parameters and did not attempt to estimate other unit-specific parameters.
IGIRF is sensitive to the choice of the guide function, and the model-specific implementation used by \citet{park20} performed better than the generic implementation of IGIRF which is currently available in \code{spatPomp}.
To our knowledge, {\ibpf} is therefore advancing the current limits for scalability of simulation-based maximum-likelihood inference for metapopulation dynamics.

@@

We compartmentalize the population of each town into susceptible ($S$), exposed ($E$), infectious ($I$), and recovered/removed ($R$) categories.
The numbers of individuals in each compartment for town $\unit$ at time $t$ are denoted by integer-valued random variables $S_\unit(t)$, $E_\unit(t)$, $I_\unit(t)$, and $R_\unit(t)$.
The population dynamics are written in terms of counting processes $N_{AB,\unit}(t)$ enumerating cumulative transitions in town $\unit$, up to time $t$, between compartments $A$ and $B$.

@@

We enumerate the $U=20$ towns studied by \citet{he10} in decreasing size, so that $\unit=1$ corresponds to London.

@@

The number of recovered individuals $R_\unit(t)$ in town $\unit$ is therefore defined implicitly.

@@

In \eqref{eq:dEdt}, $\alpha$ is a mixing exponent modeling inhomogeneous contact rates within a town, and $\iota$ models immigration of infected individuals which is appropriate when analyzing a subset of cities that cannot be treated as a closed system.
The number of travelers from town $\unit$ to $\altUnit$ is denoted by $v_{\unit\altUnit}$. 

@@

where $\dist(\unit,\altUnit)$ denotes the distance between town $\unit$ and town $\altUnit$, $P_\unit$ is the average population for town $\unit$ across time, $\bar{P}$ is the average population across towns, and $\overline{\dist}$ is the average distance between a randomly chosen pair of towns.
Here, we model $v_{\unit\altUnit}$ as fixed through time and symmetric between any two arbitrary towns, though a natural extension would allow for temporal variation and asymmetric movement between two towns.

@@

\noindent This work was supported by National Science Foundation grants DMS-1761603 and DMS 1646108.
We acknowledge an anonymous referee from a previous article who recommended that we should try implementing a block particle filter.
